---
title: "data-collection"
author: "Toby Chiu"
date: "2023-07-06"
output: html_document
---

Data is retrieved using package from [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
Underlying data is from FBref and transfermarkt (TM).

## Basic Setup

```{r installing packages}
# install.packages("worldfootballR")
library(worldfootballR)
library(tidyverse)
library(stringi)
```
## Obtaining Player Databases

Getting a list of all players

```{r}
# Database of all players
mapped_players = player_dictionary_mapping()
```

Getting stats for each player in 2022-23 EPL

```{r}
big5_player_standard <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "standard", team_or_player= "player")
big5_player_shooting <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "shooting", team_or_player= "player")
big5_player_passing <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "passing", team_or_player= "player")
big5_player_passing_types <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "passing_types", team_or_player= "player")
big5_player_gca <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "gca", team_or_player= "player")
big5_player_defense <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "defense", team_or_player= "player")
big5_player_possession <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "possession", team_or_player= "player")
big5_player_playing_time <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "playing_time", team_or_player= "player")
big5_player_misc <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "misc", team_or_player= "player")
big5_player_keepers <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "keepers", team_or_player= "player")
big5_player_keepers_adv <- fb_big5_advanced_season_stats(season_end_year= c(2019:2023), stat_type= "keepers_adv", team_or_player= "player")
```

Getting transfer valuations for all players in top 5 leagues over last 5 seasons

```{r}
player_valuations <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"), start_year = c(2019:2023))
```

## Data Cleaning

Creating a common field between FBref and transfermarkt datasets by combining the name and year to create a unique ID for each entry.

```{r}
# FBref datasets
big5_player_standard <- big5_player_standard %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_shooting <- big5_player_shooting %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_passing <- big5_player_passing %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_passing_types <- big5_player_passing_types %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_gca <- big5_player_gca %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_defense <- big5_player_defense %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_possession <- big5_player_possession %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_playing_time <- big5_player_playing_time %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_misc <- big5_player_misc %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_keepers <- big5_player_keepers %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)
big5_player_keepers_adv <- big5_player_keepers_adv %>% 
  unite("unique_ID", c("Player", "Season_End_Year"), sep="_", remove=FALSE)

# TM dataset
player_valuations <- player_valuations %>% 
  unite("unique_ID", c("player_name", "season_start_year"), sep="_", remove=FALSE)
```

We can also seperate the transfermarkt datasets into two datasets, one for outfield players and one for goalkeepers, given their widely differing skillsets and valuation techniques.

```{r}
# Splitting our TM dataset into outfielders and goalkeepers.
outfield_player_valuations <- player_valuations %>% 
  filter(player_position != "Goalkeeper")

keeper_player_valuations <- player_valuations %>% 
  filter(player_position == "Goalkeeper")
```

## Exporting Raw Datasets

We will finish the data cleaning in Pandas.


